---
title: "Theoretical Modeling"
output: html_notebook
---

Idea: 
Explore the impact of QSEQ1~5 on EQ5D

Model:
Linear Regression -> 

Problem: 
- Do not know the weights of QSEQ1~5, which are modeled off of EuroQoL research 
and fitted to Korean standards based on prior research.
- Ignoring the impact of demographic and medical qualities of patients, which 
will likely be significant

```{r}
library(tidyverse)
library(haven)
library(lme4)
library(lmerTest)  
library(sjPlot) 
library(ggplot2)
library(dplyr)
```


```{r}
df <- read_csv("NTM_KOREA_subdata.csv")

model_data <- df %>%
  select(SUBJNO, time, CMQUIT_GROUP_2, QSEQQALY,
         QSEQ1, QSEQ2, QSEQ3, QSEQ4, QSEQ5) %>%
  mutate(time = factor(time, levels = c("B", "6M", "1Y", "18M", "2Y"))) %>%
  mutate(patient_group = as.factor(CMQUIT_GROUP_2)) %>%
  select(-CMQUIT_GROUP_2)
```
```{r, warning=FALSE}
# longitudinal plot for reference
ggplot(model_data, aes(x = time, y = QSEQQALY, color = patient_group)) +
  stat_summary(fun = mean, geom = "point") +
  stat_summary(fun = mean, geom = "line", aes(group = patient_group)) +
  stat_summary(fun.data = mean_se, geom = "errorbar", 
               width = 0.2, position = position_dodge(width = .1)) +
  theme_minimal() +
  labs(title = "EQ-5D-5L Scores Over Time by Patient Group",
       y = "EQ-5D-5L Score",
       x = "Time Point")
```
# Mixed-Effects Model Structure:

Fixed effects: 
The dimension scores (QSEQ1-5), time points, and patient groups
Random effects: 
Patient-specific intercepts to account for within-subject correlation
(This structure acknowledges that measurements from the same patient are related)

```{r}
eq5d_model <- lmer(QSEQQALY ~ 
                   QSEQ1 + QSEQ2 + QSEQ3 + QSEQ4 + QSEQ5 +  # Dimension scores
                   time + patient_group +                   
                   (1|SUBJNO),                              # Random intercept for each patient
                   data = model_data)

summary(eq5d_model)
```
Model Interpretation:
> coefficients for QSEQ1-5 = how much each dimension contributes to the overall EQ-5D score
> time coefficients = show the temporal trend in scores
> patient_group coefficients = systematic differences between groups
> random effects = how much individual patients deviate from the population average


ToDo: 
- Check if the residuals are normally distributed & homoscedastic
- Look for any systematic patterns in the random effects
- Examine the VIF values for multicollinearity; Note: values > 5 suggest problems
- Consider if the linear relationship assumption holds => Appears mostly linear
```{r}
#  Residual plots
plot(eq5d_model) 
# Not homoscedastic 
# - probably b/c i'm not using the tobit model to account for the cutoff value of 1.0
# Appears somewhat normally distrubuted

# Check for multicollinearity
library(car)
vif(eq5d_model)

# random effects (= how much individual patients deviate from the population average)
ranef(eq5d_model)
```
```{r}
# Standardize predictors for comparable coefficients
model_data_scaled <- model_data %>%
  mutate(across(starts_with("QSEQ"), scale))

eq5d_model_scaled <- lmer(QSEQQALY ~ 
                         QSEQ1 + QSEQ2 + QSEQ3 + QSEQ4 + QSEQ5 +
                         time + patient_group +
                         (1|SUBJNO),
                         data = model_data_scaled)
summary(eq5d_model_scaled)
```

