---
title: "modification - binomial logistic"
author: "Youbeen Shim"
date: "2025-03-25"
output: html_document
---

Primary Question:
Using 6M & delta of EQ5D (QALY), we want to predict CMQUIT_GROUP_2 ( A&B vs. C )

Secondary Question:
What is the contribution of QALY in completing our primary goal?

Binomial Logistic Regression Model for Predicting Treatment Response in NTM-PD

```{r}
library(tidyverse)   # For data manipulation
library(caret)       # For model evaluation
library(pROC)        # For ROC curves
library(sjPlot)      # For visualization of results
library(car)         # For VIF calculation

# Load the data
data <- read_csv("a) NTM-KOREA_EQ5D5L_QOLB_18m_20250309.csv")

# Filter for baseline data only (time == "B")
baseline_data <- data %>% 
  filter(time == "B") %>%
  # Select relevant variables based on clinical importance
  select(SUBJNO, CMQUIT_GROUP_2, QSEQQALY, 
         # Demographic variables
         AGE, SEX, # DMSMKEVER, 
         # Medical history variables
         MHTBTX, MHTNTX, MHCOPD, MHASTH, 
         # Microbiological variables  
         MBEDX, # MBDXSPRYN,
         # Radiographic variables
         RFCTFD, RFCTCAVI,
         # Clinical variables
         VSBMI, LBWBC, LBALB, LBESR, LBCRP,
         # Pulmonary function
         FTPFFEVP, FTPFFVCP, FTPFFEVC, FTPFDLCO,
         # Disease severity
         BACES)  %>%
  rename(QSEQQALY_B = QSEQQALY)

updated_eq5d <- data %>%
  filter(time == "6M") %>%
  select(SUBJNO, QSEQQALY) %>%
  rename(QSEQQALY_6M = QSEQQALY)

baseline_data <- left_join(baseline_data, updated_eq5d, by = "SUBJNO") %>%
  mutate(QSEQQALY_delta = QSEQQALY_6M - QSEQQALY_B)


# Create the binary outcome variable: "Continue" (A or B) vs. "Discontinue" (C)
baseline_data <- baseline_data %>%
  mutate(treatment_status = case_when(
    CMQUIT_GROUP_2 %in% c("A", "B") ~ "Continue",
    CMQUIT_GROUP_2 == "C" ~ "Discontinue",
    TRUE ~ NA_character_
  )) %>%
  mutate(treatment_status = factor(treatment_status, levels = c("Continue", "Discontinue")))

# Check distribution of the new binary outcome
table(baseline_data$treatment_status, useNA = "ifany")

# Check for missing data
missing_data <- baseline_data %>%
  summarise(across(everything(), ~sum(is.na(.)))) %>%
  pivot_longer(cols = everything(), 
               names_to = "variable", 
               values_to = "missing_count") %>%
  filter(missing_count > 0) %>%
  arrange(desc(missing_count))

print(missing_data)

# Factorize categorical variables
baseline_data <- baseline_data %>%
  mutate(
    SEX = factor(SEX),
    # DMSMKEVER = factor(DMSMKEVER),
    MHTBTX = factor(MHTBTX),
    MHTNTX = factor(MHTNTX),
    MHCOPD = factor(MHCOPD),
    MHASTH = factor(MHASTH),
    # MBDXSPRYN = factor(MBDXSPRYN),
    RFCTFD = factor(RFCTFD),
    RFCTCAVI = factor(RFCTCAVI)
  ) %>%
  drop_na() # TODO: Should figure out how to handle NA

summary(baseline_data)
```

2. Baseline model without QSEQQALY
```{r}
baseline_model <- glm(treatment_status ~ 
                    AGE + SEX + # DMSMKEVER + 
                    MHTBTX + MHTNTX + 
                    # MBDXSPRYN +
                    RFCTFD + 
                    VSBMI + FTPFDLCO + BACES, 
                  data = baseline_data, 
                  family = binomial())

# Model summary
summary(baseline_model)

# Assess model performance
# Calculate predictions
predicted_probs_baseline <- predict(baseline_model, type = "response")
predicted_class_baseline <- ifelse(predicted_probs_baseline > 0.5, "Discontinue", "Continue")

# Create confusion matrix
confmat_baseline <- table(Predicted = predicted_class_baseline, 
                        Actual = baseline_data$treatment_status)
print(confmat_baseline)

# accuracy
accuracy_baseline <- sum(diag(confmat_baseline)) / sum(confmat_baseline)
print(paste("Baseline model accuracy:", round(accuracy_baseline * 100, 2), "%"))

# ROC and AUC for baseline model
roc_baseline <- roc(baseline_data$treatment_status, predicted_probs_baseline)
auc_baseline <- auc(roc_baseline)
print(paste("Baseline model AUC:", round(auc_baseline, 4)))

# Nagelkerke's R² for baseline model
null_model <- glm(treatment_status ~ 1, data = baseline_data, family = binomial())
r2_nagelkerke_baseline <- 1 - exp((null_model$deviance - baseline_model$deviance) / nrow(baseline_data))
r2_nagelkerke_baseline <- r2_nagelkerke_baseline / (1 - exp(-null_model$deviance / nrow(baseline_data)))
print(paste("Baseline model Nagelkerke's R²:", round(r2_nagelkerke_baseline, 4)))
```

3. Full model with QALY (6M) & delta (6M - B)
```{r}
full_model <- glm(treatment_status ~ 
                QSEQQALY_6M + QSEQQALY_delta +
                AGE + SEX + # DMSMKEVER + 
                MHTBTX + MHTNTX + 
                # MBDXSPRYN + 
                RFCTFD + 
                VSBMI + FTPFDLCO + BACES, 
              data = baseline_data, 
              family = binomial())

# Model summary
summary(full_model)

# predictions
predicted_probs_full <- predict(full_model, type = "response")
predicted_class_full <- ifelse(predicted_probs_full > 0.5, "Discontinue", "Continue")

# confusion matrix
confmat_full <- table(Predicted = predicted_class_full, 
                    Actual = baseline_data$treatment_status)
print(confmat_full)

# accuracy
accuracy_full <- sum(diag(confmat_full)) / sum(confmat_full)
print(paste("Full model accuracy:", round(accuracy_full * 100, 2), "%"))

# ROC and AUC for full model
roc_full <- roc(baseline_data$treatment_status, predicted_probs_full)
auc_full <- auc(roc_full)
print(paste("Full model AUC:", round(auc_full, 4)))

# Nagelkerke's R² for full model
r2_nagelkerke_full <- 1 - exp((null_model$deviance - full_model$deviance) / nrow(baseline_data))
r2_nagelkerke_full <- r2_nagelkerke_full / (1 - exp(-null_model$deviance / nrow(baseline_data)))
print(paste("Full model Nagelkerke's R²:", round(r2_nagelkerke_full, 4)))

# contribution of QALY + delta
r2_difference <- r2_nagelkerke_full - r2_nagelkerke_baseline
print(paste("QSEQQALY contribution to Nagelkerke's R²:", round(r2_difference, 4)))
print(paste("Percent improvement:", round(r2_difference/r2_nagelkerke_baseline * 100, 2), "%"))

# Likelihood ratio test to compare models
lr_test <- anova(baseline_model, full_model, test = "Chisq")
print(lr_test)
```

4. QSEQQALY-only model for comparison
```{r}
qseqqaly_only_model <- glm(treatment_status ~ QSEQQALY_6M + QSEQQALY_delta, 
                         data = baseline_data, 
                         family = binomial())

# Model summary
summary(qseqqaly_only_model)

# predictions
predicted_probs_qseqqaly <- predict(qseqqaly_only_model, type = "response")
predicted_class_qseqqaly <- ifelse(predicted_probs_qseqqaly > 0.5, "Discontinue", "Continue")

# confusion matrix
confmat_qseqqaly <- table(Predicted = predicted_class_qseqqaly, 
                        Actual = baseline_data$treatment_status)
print(confmat_qseqqaly)

# accuracy
accuracy_qseqqaly <- sum(diag(confmat_qseqqaly)) / sum(confmat_qseqqaly)
print(paste("QSEQQALY-only model accuracy:", round(accuracy_qseqqaly * 100, 2), "%"))

# ROC and AUC for QSEQQALY-only model
roc_qseqqaly <- roc(baseline_data$treatment_status, predicted_probs_qseqqaly)
auc_qseqqaly <- auc(roc_qseqqaly)
print(paste("QSEQQALY-only model AUC:", round(auc_qseqqaly, 4)))

# Nagelkerke's R² for QSEQQALY-only model
r2_nagelkerke_qseqqaly <- 1 - exp((null_model$deviance - qseqqaly_only_model$deviance) / nrow(baseline_data))
r2_nagelkerke_qseqqaly <- r2_nagelkerke_qseqqaly / (1 - exp(-null_model$deviance / nrow(baseline_data)))
print(paste("QSEQQALY-only model Nagelkerke's R²:", round(r2_nagelkerke_qseqqaly, 4)))
```

5. Cross-validation to evaluate robustness (skipped)

6. Variable importance analysis
```{r}
# odds ratios with confidence intervals
odds_ratios <- exp(cbind(OR = coef(full_model), confint(full_model)))
print("Odds ratios with 95% confidence intervals:")
print(odds_ratios)

# variable importance by drop-in-deviance method
calculate_importance <- function(full_model, data) {
  # Get all predictors
  predictors <- attr(terms(full_model), "term.labels")
  
  # Initialize results dataframe
  importance_df <- data.frame(
    variable = character(),
    deviance = numeric(),
    chi_square = numeric(),
    p_value = numeric(),
    importance = numeric(),
    stringsAsFactors = FALSE
  )
  
  # Full model deviance
  full_deviance <- full_model$deviance
  
  # Loop through each predictor
  for (pred in predictors) {
    # Create formula without this predictor
    reduced_formula <- as.formula(
      paste("treatment_status ~", 
            paste(predictors[predictors != pred], collapse = " + "))
    )
    
    # Fit reduced model
    reduced_model <- glm(reduced_formula, data = data, family = binomial())
    reduced_deviance <- reduced_model$deviance
    
    # Calculate chi-square
    chi_sq <- reduced_deviance - full_deviance
    df <- full_model$df.residual - reduced_model$df.residual
    p_val <- pchisq(chi_sq, df = df, lower.tail = FALSE)
    
    # Calculate normalized importance (chi-square per df)
    importance <- chi_sq / df
    
    # Add to results
    importance_df <- rbind(importance_df, data.frame(
      variable = pred,
      deviance = reduced_deviance,
      chi_square = chi_sq,
      p_value = p_val,
      importance = importance
    ))
  }
  
  # Sort by importance
  importance_df <- importance_df %>%
    arrange(desc(importance))
  
  return(importance_df)
}

# variable importance
var_importance <- calculate_importance(full_model, baseline_data)
print(var_importance)
```

7. Visualization of QSEQQALY's effect on treatment discontinuation
```{r}
# grid of QSEQQALY values covering the observed range
qseqqaly_range <- seq(min(baseline_data$QSEQQALY_6M, na.rm = TRUE), 
                     max(baseline_data$QSEQQALY_6M, na.rm = TRUE), 
                     length.out = 100)

# creating data for prediction using median/modal values for other variables
new_data <- expand.grid(
  QSEQQALY_6M = qseqqaly_range,
  QSEQQALY_delta = median(baseline_data$QSEQQALY_delta, na.rm = TRUE),
  AGE = median(baseline_data$AGE, na.rm = TRUE),
  SEX = levels(baseline_data$SEX)[1],
  # DMSMKEVER = levels(baseline_data$DMSMKEVER)[1],
  MHTBTX = levels(baseline_data$MHTBTX)[1],
  MHTNTX = levels(baseline_data$MHTNTX)[1],
  # MBDXSPRYN = levels(baseline_data$MBDXSPRYN)[1],
  RFCTFD = levels(baseline_data$RFCTFD)[1],
  VSBMI = median(baseline_data$VSBMI, na.rm = TRUE),
  FTPFDLCO = median(baseline_data$FTPFDLCO, na.rm = TRUE),
  BACES = median(baseline_data$BACES, na.rm = TRUE)
)

# predicted probabilities
new_data$predicted_prob <- predict(full_model, newdata = new_data, type = "response")

ggplot(new_data, aes(x = QSEQQALY_6M, y = predicted_prob)) +
  geom_line(size = 1.2, color = "#3366CC") +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "red") +
  labs(title = "Probability of Treatment Discontinuation by QSEQQALY Score",
       x = "QSEQQALY Score at Baseline",
       y = "Probability of Discontinuation (Group C)") +
  theme_bw()
```

8. ROC curve comparison
```{r}
# ROC curves for all models
roc_plot <- ggroc(list(
  "Baseline Model" = roc_baseline,
  "Full Model" = roc_full,
  "QSEQQALY-only Model" = roc_qseqqaly
)) +
  labs(title = "ROC Curves for Treatment Discontinuation Prediction",
       subtitle = paste("AUC - Baseline:", round(auc_baseline, 3),
                        "| Full:", round(auc_full, 3),
                        "| QSEQQALY-only:", round(auc_qseqqaly, 3))) +
  theme_bw()

print(roc_plot)
```



```{r}
# 9. Final model and summary
# Select the most important variables from var_importance
top_vars <- head(var_importance$variable, 5)  # Top 5 most important variables

# Create formula for final model
final_formula <- as.formula(paste("treatment_status ~", paste(top_vars, collapse = " + ")))

# Build final parsimonious model
final_model <- glm(final_formula, data = baseline_data, family = binomial())

# Summary of final model
summary(final_model)

# Calculate predictions
predicted_probs_final <- predict(final_model, type = "response")
predicted_class_final <- ifelse(predicted_probs_final > 0.5, "Discontinue", "Continue")

# Calculate accuracy
confmat_final <- table(Predicted = predicted_class_final, 
                     Actual = baseline_data$treatment_status)
accuracy_final <- sum(diag(confmat_final)) / sum(confmat_final)
print(paste("Final model accuracy:", round(accuracy_final * 100, 2), "%"))

# Calculate ROC and AUC for final model
roc_final <- roc(baseline_data$treatment_status, predicted_probs_final)
auc_final <- auc(roc_final)
print(paste("Final model AUC:", round(auc_final, 4)))

# Calculate Nagelkerke's R² for final model
r2_nagelkerke_final <- 1 - exp((null_model$deviance - final_model$deviance) / nrow(baseline_data))
r2_nagelkerke_final <- r2_nagelkerke_final / (1 - exp(-null_model$deviance / nrow(baseline_data)))
print(paste("Final model Nagelkerke's R²:", round(r2_nagelkerke_final, 4)))

# 10. Comprehensive model comparison
model_comparison <- data.frame(
  Model = c("Baseline (without QSEQQALY)", 
           "Full (with QSEQQALY)",
           "QSEQQALY-only",
           "Final parsimonious model"),
  Accuracy = c(accuracy_baseline, 
              accuracy_full, 
              accuracy_qseqqaly,
              accuracy_final) * 100,
  AUC = c(auc_baseline,
         auc_full,
         auc_qseqqaly,
         auc_final),
  Nagelkerke_R2 = c(r2_nagelkerke_baseline,
                   r2_nagelkerke_full,
                   r2_nagelkerke_qseqqaly,
                   r2_nagelkerke_final),
  Improvement = c(NA,
                 (r2_nagelkerke_full - r2_nagelkerke_baseline)/r2_nagelkerke_baseline * 100,
                 NA,
                 (r2_nagelkerke_final - r2_nagelkerke_baseline)/r2_nagelkerke_baseline * 100)
)

# Format the table
model_comparison$Accuracy <- round(model_comparison$Accuracy, 2)
model_comparison$AUC <- round(model_comparison$AUC, 4)
model_comparison$Nagelkerke_R2 <- round(model_comparison$Nagelkerke_R2, 4)
model_comparison$Improvement <- round(model_comparison$Improvement, 2)
model_comparison$Improvement[c(1,3)] <- NA

print(model_comparison)

# 11. QSEQQALY contribution summary
# Calculate the relative contribution of QSEQQALY to the predictive power
qseqqaly_contribution <- r2_nagelkerke_qseqqaly / r2_nagelkerke_full * 100
qseqqaly_improvement <- (r2_nagelkerke_full - r2_nagelkerke_baseline) / r2_nagelkerke_baseline * 100

qseqqaly_summary <- data.frame(
  Metric = c("QSEQQALY-only model R²",
             "Full model R²",
             "Baseline model R²",
             "QSEQQALY absolute contribution to R²",
             "QSEQQALY relative contribution to full model (%)",
             "Improvement in R² due to QSEQQALY (%)",
             "QSEQQALY variable importance rank",
             "QSEQQALY odds ratio",
             "QSEQQALY p-value"),
  Value = c(round(r2_nagelkerke_qseqqaly, 4),
            round(r2_nagelkerke_full, 4),
            round(r2_nagelkerke_baseline, 4),
            round(r2_nagelkerke_full - r2_nagelkerke_baseline, 4),
            round(qseqqaly_contribution, 2),
            round(qseqqaly_improvement, 2),
            which(var_importance$variable == "QSEQQALY_6M"),
            round(odds_ratios["QSEQQALY_6M", "OR"], 4),
            round(summary(full_model)$coefficients["QSEQQALY_6M", "Pr(>|z|)"], 4))
)

print(qseqqaly_summary)

# Final conclusion
cat("BINARY MODEL CONCLUSION:\n")
p_value <- summary(full_model)$coefficients["QSEQQALY_6M", "Pr(>|z|)"]
if (!is.na(p_value) && p_value < 0.05) {
  cat("QSEQQALY at baseline is a statistically significant predictor of treatment discontinuation.\n")
} else {
  cat("QSEQQALY at baseline is NOT a statistically significant predictor of treatment discontinuation.\n")
}

cat(paste("QSEQQALY contributes", round(qseqqaly_contribution, 1), 
          "% of the predictive power in the full model.\n"))
cat(paste("Including QSEQQALY improves the model's explanatory power by", 
          round(qseqqaly_improvement, 1), "% compared to baseline.\n"))

# Compare performance to original multinomial model
cat("\nCOMPARISON TO ORIGINAL MULTINOMIAL MODEL:\n")
cat("This binary classification approach (predicting treatment discontinuation vs. continuation) ")
cat("should provide better predictive performance than the original multinomial model ")
cat("by focusing on a clearer clinical distinction between patients who tolerate treatment ")
cat("versus those who discontinue due to adverse events.\n")
```

