---
title: "regression methods"
author: "Youbeen Shim"
date: "2025-03-21"
output: html_document
---

Primary Question:
Using BASELINE variables, we want to predict CMQUIT_GROUP_2 (which is the treatment response group)

Secondary Question:
What is the contribution of baseline QSEQQALY in completing our primary goal?

TODO - Variations
-> Using 6M                 ; regression_methods_6m.Rmd
-> Using delta (6M - B)     ; regression_methods_delta.Rmd
-> (new) Using 6M & delta   ; regression_methods_both.Rmd

-> distinguishing between A&B vs C (binomial logistic regression) ; regression_mod_binom.Rmd

-> multilevel modeling w/ age, gender, and smoking(?) (& repeat above)

Multinomial Logistic Regression Model for Predicting Treatment Response in NTM-PD

1. Load libraries and prepare data
@TODO: Handle NA's
```{r}
library(tidyverse)   
library(nnet)        # For multinomial logistic regression
library(caret)       # For model evaluation
library(pROC)        # For ROC curves
library(sjPlot)      # For visualization of results
library(car)         # For VIF calculation

# Load the data
data <- read_csv("a) NTM-KOREA_EQ5D5L_QOLB_18m_20250309.csv")

# Filter for baseline data only (time == "B")
baseline_data <- data %>% 
  filter(time == "B") %>%
  # Select relevant variables based on clinical importance
  select(SUBJNO, CMQUIT_GROUP_2, QSEQQALY, 
         # Demographic variables
         AGE, SEX, # DMSMKEVER, 
         # Medical history variables
         MHTBTX, MHTNTX, MHCOPD, MHASTH, # unnecessary 
         # Microbiological variables  
         MBEDX, # MBDXSPRYN, # unnecessary 
         # Radiographic variables
         # RFCTFD, # unnecessary 
         RFCTCAVI,
         # Clinical variables
         VSBMI, 
         LBWBC, # unnecessary 
         LBALB, # duplicate with BMI
         LBESR, 
         LBCRP, # duplicate with ESR
         # Pulmonary function
         FTPFFEVP, FTPFFVCP, FTPFFEVC, FTPFDLCO, # unnecessary 
         # Disease severity
         BACES) # BACES includes age/bmi/cavi/esr - with and without 

# Check for missing data
missing_data <- baseline_data %>%
  summarise(across(everything(), ~sum(is.na(.)))) %>%
  pivot_longer(cols = everything(), 
               names_to = "variable", 
               values_to = "missing_count") %>%
  filter(missing_count > 0) %>%
  arrange(desc(missing_count))

print(missing_data)

# Factorize categorical variables
baseline_data <- baseline_data %>%
  mutate(
    CMQUIT_GROUP_2 = factor(CMQUIT_GROUP_2, levels = c("A", "B", "C")),
    SEX = factor(SEX),
    # DMSMKEVER = factor(DMSMKEVER),
    MHTBTX = factor(MHTBTX),
    MHTNTX = factor(MHTNTX),
    MHCOPD = factor(MHCOPD),
    MHASTH = factor(MHASTH),
    # MBDXSPRYN = factor(MBDXSPRYN),
    RFCTFD = factor(RFCTFD),
    RFCTCAVI = factor(RFCTCAVI)
  ) %>%
  drop_na() # TODO: Should figure out how to handle NA


summary(baseline_data)
```

```{r}
# Age, Sex, BMI 

b_data <- data %>% filter(time == 'B')
hist(b_data$AGE)

table(b_data$SEX)
table(b_data$SEX, b_data$CMQUIT_GROUP_2)

hist(b_data$VSBMI)

hist(b_data$BACES) # 0&1, 2&3, 4&5 
```


2. Baseline model without QSEQQALY
```{r}
# Create a baseline model without QSEQQALY 
baseline_model <- multinom(CMQUIT_GROUP_2 ~ 
                          AGE + SEX + #DMSMKEVER + 
                          MHTBTX + MHTNTX + 
                          #MBDXSPRYN + 
                          RFCTFD + 
                          VSBMI + FTPFDLCO + BACES, 
                        data = baseline_data, 
                        trace = FALSE)

# Model summary
summary(baseline_model)

# Examine model fit
baseline_fit <- fitted(baseline_model)

# Calculate accuracy
predicted_classes_baseline <- predict(baseline_model, type = "class")
accuracy_baseline <- mean(predicted_classes_baseline == baseline_data$CMQUIT_GROUP_2, na.rm = TRUE)
print(paste("Baseline model accuracy:", round(accuracy_baseline * 100, 2), "%"))

# Create confusion matrix
confmat_baseline <- table(Predicted = predicted_classes_baseline, 
                        Actual = baseline_data$CMQUIT_GROUP_2)
print(confmat_baseline)

# Calculate pseudo R-squared
null_model <- multinom(CMQUIT_GROUP_2 ~ 1, data = baseline_data, trace = FALSE)
mcfadden_r2_baseline <- 1 - logLik(baseline_model)/logLik(null_model)
print(paste("McFadden's pseudo R-squared (baseline model):", round(as.numeric(mcfadden_r2_baseline), 4)))
```

3. QSEQQALY-only model for comparison
```{r}
# Create a model with QSEQQALY only
qseqqaly_only_model <- multinom(CMQUIT_GROUP_2 ~ QSEQQALY, 
                              data = baseline_data, 
                              trace = FALSE)

# Calculate McFadden's pseudo R-squared
mcfadden_r2_qseqqaly <- 1 - logLik(qseqqaly_only_model)/logLik(null_model)
print(paste("McFadden's pseudo R-squared (QSEQQALY-only model):", round(as.numeric(mcfadden_r2_qseqqaly), 4)))

# Calculate accuracy
predicted_classes_qseqqaly <- predict(qseqqaly_only_model, type = "class")
accuracy_qseqqaly <- mean(predicted_classes_qseqqaly == baseline_data$CMQUIT_GROUP_2, na.rm = TRUE)
print(paste("QSEQQALY-only model accuracy:", round(accuracy_qseqqaly * 100, 2), "%"))
```


4. Full model with QSEQQALY
```{r}
# Create full model including QSEQQALY
full_model <- multinom(CMQUIT_GROUP_2 ~ 
                       QSEQQALY +
                       AGE + SEX + #DMSMKEVER + 
                       MHTBTX + MHTNTX + 
                       # MBDXSPRYN + 
                       RFCTFD + 
                       VSBMI + FTPFDLCO + BACES, 
                     data = baseline_data, 
                     trace = FALSE)

# Model summary
summary(full_model)

# Examine model fit
full_fit <- fitted(full_model)

# Calculate accuracy
predicted_classes_full <- predict(full_model, type = "class")
accuracy_full <- mean(predicted_classes_full == baseline_data$CMQUIT_GROUP_2, na.rm = TRUE)
print(paste("Full model accuracy:", round(accuracy_full * 100, 2), "%"))

# Create confusion matrix
confmat_full <- table(Predicted = predicted_classes_full, 
                     Actual = baseline_data$CMQUIT_GROUP_2)
print(confmat_full)

# Calculate McFadden's pseudo R-squared
mcfadden_r2_full <- 1 - logLik(full_model)/logLik(null_model)
print(paste("McFadden's pseudo R-squared (full model):", round(as.numeric(mcfadden_r2_full), 4)))

# Calculate the contribution of QSEQQALY
r2_difference <- as.numeric(mcfadden_r2_full) - as.numeric(mcfadden_r2_baseline)
print(paste("QSEQQALY contribution to McFadden's R-squared:", round(r2_difference, 4)))
print(paste("Percent improvement:", round(r2_difference/as.numeric(mcfadden_r2_baseline) * 100, 2), "%"))

# Likelihood ratio test to compare models
lr_test <- anova(baseline_model, full_model)
print(lr_test)
```

5. (Extra; currently broken) Cross-validation to evaluate robustness
```{r}
# Set up cross-validation
set.seed(123)
train_control <- trainControl(
  method = "cv",
  number = 5,
  classProbs = TRUE,
  summaryFunction = multiClassSummary
)

# Convert variables to numeric for caret
model_data <- baseline_data %>%
  mutate_if(is.factor, as.numeric)

# Train model with baseline variables only
cv_model_baseline <- train(
  #CMQUIT_GROUP_2 ~ AGE + SEX + DMSMKEVER + MHTBTX + MHTNTX + MBDXSPRYN + RFCTFD + VSBMI + FTPFDLCO + BACES,
  CMQUIT_GROUP_2 ~ AGE + SEX + MHTBTX + MHTNTX + RFCTFD + VSBMI + FTPFDLCO + BACES,
  data = model_data,
  method = "cforest",
  trControl = train_control,
  trace = FALSE
)

# Train model with QSEQQALY included
cv_model_full <- train(
  #CMQUIT_GROUP_2 ~ QSEQQALY + AGE + SEX + DMSMKEVER + MHTBTX + MHTNTX + MBDXSPRYN + RFCTFD + VSBMI + FTPFDLCO + BACES,
  CMQUIT_GROUP_2 ~ QSEQQALY + AGE + SEX + MHTBTX + MHTNTX + RFCTFD + VSBMI + FTPFDLCO + BACES,
  data = model_data,
  method = "cforest",
  trControl = train_control,
  trace = FALSE
)

# Compare results
print("Cross-validation results for baseline model:")
print(cv_model_baseline)
print(cv_model_baseline$results)

print("Cross-validation results for full model (with QSEQQALY):")
print(cv_model_full)
print(cv_model_full$results)

# Calculate improvement in cross-validated accuracy
cv_improvement <- cv_model_full$results$Accuracy - cv_model_baseline$results$Accuracy
print(paste("Cross-validated accuracy improvement with QSEQQALY:", round(cv_improvement, 4)))
```

6. Variable importance and visualization
```{r}
# Calculate odds ratios for the full model
odds_ratios <- exp(coef(full_model))
print("Odds ratios for full model:")
print(odds_ratios)

# Calculate variable importance by comparing full model to models with each variable removed
calculate_importance <- function(full_model, data) {
  # Get all predictors
  predictors <- attr(terms(full_model), "term.labels")
  
  # Initialize results dataframe
  importance_df <- data.frame(
    variable = character(),
    log_likelihood = numeric(),
    chi_square = numeric(),
    p_value = numeric(),
    importance = numeric(),
    stringsAsFactors = FALSE
  )
  
  # Full model log-likelihood
  full_ll <- logLik(full_model)
  
  # Loop through each predictor
  for (pred in predictors) {
    # Create formula without this predictor
    reduced_formula <- as.formula(
      paste("CMQUIT_GROUP_2 ~", 
            paste(predictors[predictors != pred], collapse = " + "))
    )
    
    # Fit reduced model
    reduced_model <- multinom(reduced_formula, data = data, trace = FALSE)
    reduced_ll <- logLik(reduced_model)
    
    # Calculate chi-square
    chi_sq <- -2 * (reduced_ll - full_ll)
    df <- attr(full_ll, "df") - attr(reduced_ll, "df")
    p_val <- pchisq(chi_sq, df = df, lower.tail = FALSE)
    
    # Calculate normalized importance
    importance <- as.numeric(chi_sq / df)
    
    # Add to results
    importance_df <- rbind(importance_df, data.frame(
      variable = pred,
      log_likelihood = as.numeric(reduced_ll),
      chi_square = as.numeric(chi_sq),
      p_value = p_val,
      importance = importance
    ))
  }
  
  # Sort by importance
  importance_df <- importance_df %>%
    arrange(desc(importance))
  
  return(importance_df)
}

# Calculate variable importance
var_importance <- calculate_importance(full_model, baseline_data)
print(var_importance)
```

7. Plot predicted probabilities against QSEQQALY
```{r}
# Create a grid of QSEQQALY values
qseqqaly_range <- seq(min(baseline_data$QSEQQALY, na.rm = TRUE), 
                      max(baseline_data$QSEQQALY, na.rm = TRUE), 
                      length.out = 100)

# Create a dataframe with the mean/median of other predictors
new_data <- expand.grid(
  QSEQQALY = qseqqaly_range,
  AGE = median(baseline_data$AGE, na.rm = TRUE),
  SEX = levels(baseline_data$SEX)[1],
  # DMSMKEVER = levels(baseline_data$DMSMKEVER)[1],
  MHTBTX = levels(baseline_data$MHTBTX)[1],
  MHTNTX = levels(baseline_data$MHTNTX)[1],
  # MBDXSPRYN = levels(baseline_data$MBDXSPRYN)[1],
  RFCTFD = levels(baseline_data$RFCTFD)[1],
  VSBMI = median(baseline_data$VSBMI, na.rm = TRUE),
  FTPFDLCO = median(baseline_data$FTPFDLCO, na.rm = TRUE),
  BACES = median(baseline_data$BACES, na.rm = TRUE)
)

# Get predicted probabilities
predicted_probs <- predict(full_model, newdata = new_data, type = "probs")
predicted_data <- cbind(new_data, predicted_probs)

# Convert to long format for plotting
predicted_data_long <- predicted_data %>%
  pivot_longer(cols = c(A, B, C), 
               names_to = "CMQUIT_GROUP_2", 
               values_to = "probability")

# Plot
ggplot(predicted_data_long, aes(x = QSEQQALY, y = probability, color = CMQUIT_GROUP_2)) +
  geom_line(size = 1.2) +
  labs(title = "Predicted Probabilities by QSEQQALY Score",
       x = "QSEQQALY Score at Baseline",
       y = "Predicted Probability",
       color = "Treatment Response Group") +
  scale_color_manual(values = c("A" = "#EEAAAA", "B" = "#AAAAEE", "C" = "#AAEEAA"),
                     labels = c("A" = "Culture Conversion & Completed",
                               "B" = "Failed Culture Conversion",
                               "C" = "Discontinued (Adverse Events)")) +
  theme_bw() +
  theme(legend.position = "bottom")
```

8. Final model and summary
```{r}
# Create a more parsimonious final model based on important variables
# (Include top variables identified from importance analysis)
final_model <- multinom(CMQUIT_GROUP_2 ~ 
                        QSEQQALY + FTPFDLCO + RFCTFD + MHTNTX + AGE,
                      data = baseline_data, 
                      trace = FALSE)

# Calculate accuracy
predicted_classes_final <- predict(final_model, type = "class")
accuracy_final <- mean(predicted_classes_final == baseline_data$CMQUIT_GROUP_2, na.rm = TRUE)
print(paste("Final model accuracy:", round(accuracy_final * 100, 2), "%"))

# Calculate McFadden's pseudo R-squared
mcfadden_r2_final <- 1 - logLik(final_model)/logLik(null_model)
print(paste("McFadden's pseudo R-squared (final model):", round(as.numeric(mcfadden_r2_final), 4)))

# Create a summary table of all models
model_summary <- data.frame(
  Model = c("Baseline (without QSEQQALY)", 
            "Full (with QSEQQALY)", 
            "QSEQQALY only",
            "Final parsimonious model"),
  Accuracy = c(accuracy_baseline, 
               accuracy_full, 
               accuracy_qseqqaly,
               accuracy_final) * 100,
  McFadden_R2 = c(as.numeric(mcfadden_r2_baseline),
                 as.numeric(mcfadden_r2_full),
                 as.numeric(mcfadden_r2_qseqqaly),
                 as.numeric(mcfadden_r2_final)),
  Improvement = c(NA,
                 (as.numeric(mcfadden_r2_full) - as.numeric(mcfadden_r2_baseline))/as.numeric(mcfadden_r2_baseline) * 100,
                 NA,
                 (as.numeric(mcfadden_r2_final) - as.numeric(mcfadden_r2_baseline))/as.numeric(mcfadden_r2_baseline) * 100)
)

# Format the table
model_summary$Accuracy <- round(model_summary$Accuracy, 2)
model_summary$McFadden_R2 <- round(model_summary$McFadden_R2, 4)
model_summary$Improvement <- round(model_summary$Improvement, 2)
model_summary$Improvement[c(1,3)] <- NA

print(model_summary)
```

9. Conclusions about QSEQQALY contribution
```{r}
# Calculate the relative contribution of QSEQQALY to the predictive power
qseqqaly_contribution <- as.numeric(mcfadden_r2_qseqqaly) / as.numeric(mcfadden_r2_full) * 100
qseqqaly_improvement <- (as.numeric(mcfadden_r2_full) - as.numeric(mcfadden_r2_baseline)) / as.numeric(mcfadden_r2_baseline) * 100

# Create a summary specifically about QSEQQALY
qseqqaly_summary <- data.frame(
  Metric = c("QSEQQALY-only model R²",
             "Full model R²",
             "Baseline model R²",
             "QSEQQALY absolute contribution to R²",
             "QSEQQALY relative contribution to full model (%)",
             "Improvement in R² due to QSEQQALY (%)",
             "QSEQQALY variable importance rank",
             "QSEQQALY p-value"),
  Value = c(round(as.numeric(mcfadden_r2_qseqqaly), 4),
            round(as.numeric(mcfadden_r2_full), 4),
            round(as.numeric(mcfadden_r2_baseline), 4),
            round(as.numeric(mcfadden_r2_full) - as.numeric(mcfadden_r2_baseline), 4),
            round(qseqqaly_contribution, 2),
            round(qseqqaly_improvement, 2),
            which(var_importance$variable == "QSEQQALY"),
            round(var_importance$p_value[var_importance$variable == "QSEQQALY"], 4))
)

print(qseqqaly_summary)
```

