---
title: "treatment decision making"
author: "Youbeen Shim"
date: "2025-03-21"
output: html_document
---

```{r}
library(tidyverse)
library(haven)
library(lme4)
library(lmerTest)  
library(sjPlot) 
library(ggplot2)
library(dplyr)
```

```{r}
data <- read_csv("a) NTM-KOREA_EQ5D5L_QOLB_18m_20250309.csv")
# A: 77  B: 62  C: 47 = total 186 명
table(data$CMQUIT_GROUP_2)
table(data$time)
```

```{r}
# Key Variables
# MBSPECIES + AGE + SEX + DMSMKEVER + MHTBTX + MHTNTX + MHCOPD + MHASTH + MHILD + MHBE + MHCPA + MHLCA + MHDM + MHRA + MHCLD + MHCKD + MHCV + MHMAL + MBDXSPRYN + MBEDX + MBDXCO + MBDXAVC + MBDXICL + MBDXASA + MBDXASM + MBDXOTH + RFCTFD + RFCTCAVI + VSWEIGHT + VSBMI + LBWBC + LBALB + LBESR + LBCRP + FTPFFEV + FTPFFEVP + FTPFFVC + FTPFFVCP + FTPFFEVC + FTPFDLCO + FT6MWT + BACES + CMQUIT_GROUP_2 + MBDXCO2
```


LOESS Regression, Longitudinal Analysis of Quality of Life
```{r}
# Prepare data - filter out rows with missing values for duration or quality of life score
data_filtered <- data %>%
  filter(!is.na(QSEQDUR_diff) & !is.na(QSEQQALY))

# LOESS regression for each treatment response group
loess_results <- data_filtered %>%
  group_by(CMQUIT_GROUP_2) %>%
  do({
    # Fit LOESS model for current group
    loess_fit <- loess(QSEQQALY ~ QSEQDUR_diff, data = ., 
                      span = 0.75)  # Adjust span parameter as needed for smoothness
    # Create sequence of time points for prediction
    time_seq <- seq(min(.$QSEQDUR_diff), max(.$QSEQDUR_diff), length.out = 100)
    
    # predictions (with standard errors)
    pred_vals <- predict(loess_fit, 
                        newdata = data.frame(QSEQDUR_diff = time_seq), 
                        se = TRUE)
    # predictions & 95% CI
    data.frame(
      QSEQDUR_diff = time_seq,
      loess_fit = pred_vals$fit,
      lower_bound = pred_vals$fit - 1.96 * pred_vals$se.fit,  # 95% CI lower bound
      upper_bound = pred_vals$fit + 1.96 * pred_vals$se.fit   # 95% CI upper bound
    )
  }) %>%
  ungroup()


# Create visualization of LOESS regression results
loess_plot <- ggplot() +
  # regression lines for each group
  geom_line(data = loess_results, 
            aes(x = QSEQDUR_diff/365, y = loess_fit, color = CMQUIT_GROUP_2), 
            linewidth = 1) +
  # confidence interval ribbons
  geom_ribbon(data = loess_results, 
              aes(x = QSEQDUR_diff/365, 
                  ymin = lower_bound, 
                  ymax = upper_bound, 
                  fill = CMQUIT_GROUP_2), 
              alpha = 0.2) +
  labs(title = "Health-Related Quality of Life Trends by Treatment Response Group", 
       x = "Time (years)", 
       y = "QALY weight",
       color = "Response Group",
       fill = "Response Group") +
  # Set y-axis limits and break to focus on relevant range
  # scale_y_continuous(limits = c(0, 1)) +
  scale_y_continuous(limits = c(0.7, 1)) +
  # other options
  scale_fill_manual(values = group_colors) +
  scale_color_manual(values = group_colors) +
  theme_classic() +
  theme(
    legend.position = "bottom",
    plot.title = element_text(face = "bold", size = 14),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(size = 10)
  )

print(loess_plot)
```

Tobit Regression, QALY data with upper limit of 1

Note: two errors
1) Ignoring data structure: treated observations as independent despite having 4 measurements per individual over time (baseline, 6M, 1Y, 18M), violating independence assumptions in regression
2) Separate models by group: fit separate Tobit models for each treatment group. failed to pool information across groups and didn't account for within-subject correlation.
```{r}
library(AER) 
# Tobit regression for QALY data with upper limit of 1

# Filter data for analysis
tobit_data <- data %>%
  filter(!is.na(QSEQDUR_diff) & !is.na(QSEQQALY)) %>%
  # Create normalized time variable (in years)
  mutate(time_years = QSEQDUR_diff/365)

# Basic Tobit model for each group separately
# Group A
tobit_A <- tobit(QSEQQALY ~ time_years, data = subset(tobit_data, CMQUIT_GROUP_2 == "A"), 
                 left = 0, right = 1)

# Group B
tobit_B <- tobit(QSEQQALY ~ time_years, data = subset(tobit_data, CMQUIT_GROUP_2 == "B"), 
                 left = 0, right = 1)

# Group C
tobit_C <- tobit(QSEQQALY ~ time_years, data = subset(tobit_data, CMQUIT_GROUP_2 == "C"), 
                 left = 0, right = 1)

# Combined model with interaction terms
tobit_combined <- tobit(QSEQQALY ~ time_years * CMQUIT_GROUP_2, 
                        data = tobit_data, 
                        left = 0, right = 1)

# model summaries
summary(tobit_A)
summary(tobit_B)
summary(tobit_C)
summary(tobit_combined)
```
Generate predictions from Tobit models (for visualization)
```{r}
# Function to generate predictions
predict_tobit <- function(model, newdata) {
  # Get predicted values (these will be the latent variable predictions)
  pred <- predict(model, newdata = newdata, type = "response")
  
  # Apply censoring manually to match Tobit assumptions
  pred[pred < 0] <- 0
  pred[pred > 1] <- 1
  
  return(pred)
}

# Create sequence of time points for predictions
time_seq <- seq(min(tobit_data$time_years, na.rm = TRUE), 
                max(tobit_data$time_years, na.rm = TRUE), 
                length.out = 100)

# Generate predictions for each group
# Group A
pred_data_A <- data.frame(time_years = time_seq)
pred_data_A$pred_QSEQQALY <- predict_tobit(tobit_A, newdata = pred_data_A)
pred_data_A$CMQUIT_GROUP_2 <- "A"

# Group B
pred_data_B <- data.frame(time_years = time_seq)
pred_data_B$pred_QSEQQALY <- predict_tobit(tobit_B, newdata = pred_data_B)
pred_data_B$CMQUIT_GROUP_2 <- "B"

# Group C
pred_data_C <- data.frame(time_years = time_seq)
pred_data_C$pred_QSEQQALY <- predict_tobit(tobit_C, newdata = pred_data_C)
pred_data_C$CMQUIT_GROUP_2 <- "C"

# Combine predictions
tobit_predictions <- rbind(pred_data_A, pred_data_B, pred_data_C)
```
Tobit regression results
```{r}
group_colors <- c("A" = "#EEAAAA", "B" = "#AAAAEE", "C" = "#AAEEAA")

# Tobit regression results
ggplot() +
  # raw data points
  geom_point(data = tobit_data, 
             aes(x = time_years, y = QSEQQALY, color = CMQUIT_GROUP_2), 
             alpha = 0.3, size = 1) +
  # Tobit regression lines
  geom_line(data = tobit_predictions, 
            aes(x = time_years, y = pred_QSEQQALY, color = CMQUIT_GROUP_2), 
            size = 1) +
  # Styling
  scale_color_manual(values = group_colors) +
  labs(title = "Tobit Regression of Health-Related Quality of Life by Treatment Group", 
       subtitle = "Accounting for censoring at QALY = 1",
       x = "Time (years)", 
       y = "QALY weight") +
  #scale_y_continuous(limits = c(0, 1)) +
  scale_y_continuous(limits = c(0.7, 1)) +
  theme_classic() +
  theme(
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 9),
    legend.position = "bottom"
  )
```

No Tobit models that account for multilevel modeling. (or there were and I could not implement it)

Solution:
Mixed-effects model for longitudinal data. The new code properly accounts for repeated measures by including random effects for subjects

The code,
lmm_transformed <- lmer(QSEQQALY_logit ~ time_years * CMQUIT_GROUP_2 + (1 | SUBJNO), data = tobit_data)
1) acknowledges that observations from the same subject are correlated

we also have a
2) unified model with interactions
Instead of separate models, we use a single model with interaction terms, allowing for proper statistical comparison between groups while maintaining more statistical power.

HOWEVER, 
3) we are "mimicking" Tobit using a two-step approach:
  a) Logit transformation: QSEQQALY_logit = log(QSEQQALY_bounded/(1-QSEQQALY_bounded))
  b) Group-specific minimum handling: Using actual observed minimum values per group
  c) Bounded predictions: Ensuring values stay within [0,1] range
  
The method in (3) is different from Tobit, which uses the maximum likelihood estimation to handle values that is close to the limit. Instead, we apply a logit transformation to the data, mapping [0, 1] to [-Inf, Inf] and then mapping it back to [0, 1] once prediction is complete. 

NOTE: Why are there no values between 0.9 and 1.0???
```{r}
# Load necessary libraries
library(tidyverse)
library(lme4)
library(lmerTest)
library(ggplot2)

# Prepare data for analysis - improved NA handling
tobit_data <- data %>%
  # Explicitly filter out NA values for QSEQQALY
  filter(!is.na(QSEQDUR_diff) & !is.na(QSEQQALY)) %>%
  # Create normalized time variable (in years)
  mutate(time_years = QSEQDUR_diff/365) %>%
  # Make sure SUBJNO is treated as a factor (for random effects)
  mutate(SUBJNO = factor(SUBJNO))

# Get minimum values per group (to ensure we're using accurate bounds)
min_values <- tobit_data %>%
  group_by(CMQUIT_GROUP_2) %>%
  summarize(min_QSEQQALY = min(QSEQQALY, na.rm = TRUE))

# Print minimum values to confirm
print(min_values)

# Now apply transformations with group-specific minimum values
tobit_data <- tobit_data %>%
  left_join(min_values, by = "CMQUIT_GROUP_2") %>%
  # Transform QSEQQALY values to handle censoring
  # Using a small buffer from actual minimum values
  mutate(
    min_buffer = min_QSEQQALY * 0.9, # 90% of min value as buffer
    QSEQQALY_adj = case_when(
      QSEQQALY >= 1 ~ 0.999,  # Value close to 1 for right-censored
      TRUE ~ pmax(QSEQQALY, min_buffer) # Ensure no values below group min
    ),
    # Logit transformation to handle bounds
    QSEQQALY_logit = log(QSEQQALY_adj/(1-QSEQQALY_adj))
  )

# Fit linear mixed model on transformed data
lmm_transformed <- lmer(
  QSEQQALY_logit ~ time_years * CMQUIT_GROUP_2 + (1 | SUBJNO),
  data = tobit_data,
  control = lmerControl(optimizer = "bobyqa") # More robust optimizer
)

# Model summary
model_summary <- summary(lmm_transformed)
print(model_summary)

# Create prediction data frame for plotting
# Use actual observed time range for each group
time_ranges <- tobit_data %>%
  group_by(CMQUIT_GROUP_2) %>%
  summarize(
    min_time = min(time_years, na.rm = TRUE),
    max_time = max(time_years, na.rm = TRUE)
  )

# Create custom prediction grid for all groups with group-specific time ranges
pred_grid <- do.call(rbind, lapply(unique(tobit_data$CMQUIT_GROUP_2), function(group) {
  group_range <- time_ranges %>% filter(CMQUIT_GROUP_2 == group)
  data.frame(
    time_years = seq(group_range$min_time, group_range$max_time, length.out = 100),
    CMQUIT_GROUP_2 = group
  )
}))

# Add random effect values (set to 0 for population-level predictions)
pred_grid$SUBJNO <- NA

# Get fixed effects matrix
X <- model.matrix(~ time_years * CMQUIT_GROUP_2, data = pred_grid)

# Get fixed effects coefficients
beta <- fixef(lmm_transformed)

# Compute predictions on the logit scale
pred_grid$predicted_logit <- as.vector(X %*% beta)

# Back-transform to original scale
pred_grid$predicted_QSEQQALY <- 1 / (1 + exp(-pred_grid$predicted_logit))

# Define group colors to match reference image
group_colors <- c("A" = "#EEAAAA", "B" = "#AAAAEE", "C" = "#AAEEAA")

# Create the plot with improved aesthetics
ggplot() +
  # Raw data points
  geom_point(data = tobit_data, 
             aes(x = time_years, y = QSEQQALY, color = ›), 
             alpha = 0.3, size = 1) +
  # Regression lines
  geom_line(data = pred_grid, 
            aes(x = time_years, y = predicted_QSEQQALY, color = CMQUIT_GROUP_2), 
            size = 1) +
  # Styling
  scale_color_manual(values = group_colors) +
  labs(title = "Tobit Regression of Health-Related Quality of Life by Treatment Group", 
       subtitle = "Accounting for censoring at QALY = 1",
       x = "Time (years)", 
       y = "QALY weight") +
  scale_y_continuous(limits = c(0.7, 1)) +
  theme_classic() +
  theme(
    legend.title = element_blank(),
    legend.text = element_text(size = 9),
    legend.position = "bottom"
  )
```

```{r}
ggplot() +
  # Raw data points
  geom_point(data = tobit_data, 
             aes(x = time_years, y = QSEQQALY, color = CMQUIT_GROUP_2), 
             alpha = 0.3, size = 1) +
  theme_minimal()
```

