---
title: "tobit_model_complete_rundown"
author: "Youbeen Shim"
date: "2025-05-01"
output: html_document
---

```{r}
library(tidyverse)
# library(AER)  # tobit 
library(VGAM) # tobit, alt
library(survival) # 
library(pscl) # zero-inflated models
library(ggplot2)
library(car) # VIF
library(haven)
library(lme4)
library(lmerTest)  
library(sjPlot) 
library(dplyr)

data <- read_csv("a) NTM-KOREA_EQ5D5L_QOLB_18m_20250309.csv")
# A: 77  B: 62  C: 47 = total 186 명
table(data$CMQUIT_GROUP_2)
table(data$time)


ggplot(data, aes(x = QSEQQALY)) +
  geom_histogram(bins = 30) +
  labs(title = "Distribution of QSEQQALY")

# Check for clustering at boundaries (0 and 1)
sum(data$QSEQQALY == 0) / nrow(data) * 100  # percentage at 0
sum(data$QSEQQALY == 1) / nrow(data) * 100  # percentage at 1
sum(data$QSEQQALY >= 0.9 & data$QSEQQALY < 1) / nrow(data) * 100  # percentage at 0.9-0.999

# Explore QSEQQALY by grouping variable
ggplot(data, aes(x = CMQUIT_GROUP_2, y = QSEQQALY)) +
  geom_boxplot() +
  labs(title = "QSEQQALY by Group")

# Correlation matrix of predictors
cor_matrix <- data %>%
  select(AGE, RFCTCAVI, VSBMI, LBESR) %>%
  cor(use = "complete.obs")
print(cor_matrix)


```

so the data above has something like a half of a normal curve, and then an aggressively aggregated 1 value. 
we knew this, and we want to apply the censoring logic (there are probably values beyond 1)

so we think that the tobit model is appropriate, but we want to check which variables to build the tobit 
model with

NOTE: all of the below modeling is done when time == B, i.e. only for the initial screening
```{r}
data <- data %>%
  mutate(
    CMQUIT_GROUP_2 = factor(CMQUIT_GROUP_2, levels = c("A", "B", "C")),
    SEX = factor(SEX),
    DMSMK = factor(DMSMK),
    RFCTCAVI = factor(RFCTCAVI),
    MBSPECIES = factor(MBSPECIES)
  )

# fit a separate!! tobit model with a single predictor variable
# each model shows how one variable alone relateds to QALY - Also proxy to variable importance
tobit_age <- AER::tobit(QSEQQALY ~ AGE, data = subset(data, time == "B"), left = 0, right = 1)
tobit_sex <- AER::tobit(QSEQQALY ~ SEX, data = subset(data, time == "B"), left = 0, right = 1)
tobit_smoke <- AER::tobit(QSEQQALY ~ DMSMK, data = subset(data, time == "B"), left = 0, right = 1)
tobit_cavi <- AER::tobit(QSEQQALY ~ RFCTCAVI, data = subset(data, time == "B"), left = 0, right = 1)
tobit_bmi <- AER::tobit(QSEQQALY ~ VSBMI, data = subset(data, time == "B"), left = 0, right = 1)
tobit_lbesr <- AER::tobit(QSEQQALY ~ LBESR, data = subset(data, time == "B"), left = 0, right = 1)
tobit_group <- AER::tobit(QSEQQALY ~ CMQUIT_GROUP_2, data = subset(data, time == "B"), left = 0, right = 1)
tobit_mbspecies <- AER::tobit(QSEQQALY ~ MBSPECIES, data = subset(data, time == "B"), left = 0, right = 1)

# For categorical variables, we need to get the overall significance rather than individual level p-values. 
# we use likelihood ratio tests.

null_model <- AER::tobit(QSEQQALY ~ 1, data = subset(data, time == "B"), left = 0, right = 1)

# Get p-values for categorical variables
p_sex <- anova(null_model, tobit_sex, test = "Chisq")
p_smoke <- anova(null_model, tobit_smoke, test = "Chisq")
p_cavi <- anova(null_model, tobit_cavi, test = "Chisq")
p_group <- anova(null_model, tobit_group, test = "Chisq")
# p_mbspecies <- anova(null_model, tobit_mbspecies, test = "Chisq")

# p-values and AIC to compare
summary_stats <- data.frame(
  Variable = c("AGE", "SEX", "DMSMK", "RFCTCAVI", "VSBMI", "LBESR", "GROUP"), #"SPECIES"),
  P_value = c(
    summary(tobit_age)$coefficients[2,4],
    p_sex$`Pr(>Chi)`[2],
    p_smoke$`Pr(>Chi)`[2],
    p_cavi$`Pr(>Chi)`[2],
    summary(tobit_bmi)$coefficients[2,4],
    summary(tobit_lbesr)$coefficients[2,4],
    p_group$`Pr(>Chi)`[2]
    # , p_mbspecies$`Pr(>Chi)`[2]
  ),
  AIC = c(
    AIC(tobit_age), # for model fit (lower AIC = more desireable variable)
    AIC(tobit_sex),
    AIC(tobit_smoke),
    AIC(tobit_cavi),
    AIC(tobit_bmi),
    AIC(tobit_lbesr),
    AIC(tobit_group)
    # , AIC(tobit_mbspecies) 
  )
)

# Sort by AIC
summary_stats %>% arrange(AIC)
```
Looking at above, RFCTCAVI (p=0.015) and LBESR (p=0.008) are statistically significant at the conventional 0.05 level
All other variables, including GROUP, are not statistically significant in univariate models

p-values here indicate stronger statistical significance, cavity size being the most significant predictor.

Note that a major limitation here is that it only shows univariate level significance for variables that are
not standardized/normalized. 

Below, I am intending to separate the distribution of QALY scores by both time and group assignment. 
Notably, while most of the shape remains consistent, for the treatment discontinued group, the rate at which the
calculation comes out to 0.9 and 1.0 vary between time periods. 
```{r}
data <- data %>%
  mutate(
    time_factor = factor(time, levels = c("B", "6M", "1Y", "18M")),
    time_numeric = case_when(
      time == "B" ~ 0,
      time == "6M" ~ 0.5,
      time == "1Y" ~ 1,
      time == "18M" ~ 1.5
    ),
    QSEQQALY_adj = ifelse(QSEQQALY >= 1, 0.9999, QSEQQALY)
  )

ggplot(data, aes(x = QSEQQALY, fill = CMQUIT_GROUP_2)) +
  geom_histogram(bins = 20, alpha = 0.7) +
  facet_grid(time_factor ~ CMQUIT_GROUP_2) +
  labs(title = "Distribution of QALY by Time and Group")

# Check for balancedness in your design
table(data$CMQUIT_GROUP_2, data$time_factor)
```
Modeling issue
Tobit modeling is not compatible with 


Data quality issue


```{r}
b_data <- data %>% filter(time == "B")

table(b_data$QSEQQALY, b_data$CMQUIT_GROUP_2)
table(b_data$CMQUIT_GROUP_2)
```

```{r}
data_check <- data %>%
  select(QSEQQALY, time, AGE, SEX, VSBMI, LBESR, DMSMK, RFCTCAVI, MBSPECIES, CMQUIT_GROUP_2, SUBJNO)

na_count_b <- sapply(data_check %>% filter(time == "B"), function(y) sum(length(which(is.na(y)))))
na_count_6m <- sapply(data_check %>% filter(time == "6M"), function(y) sum(length(which(is.na(y)))))
na_count_1y <- sapply(data_check %>% filter(time == "1Y"), function(y) sum(length(which(is.na(y)))))
na_count_18m <- sapply(data_check %>% filter(time == "18M"), function(y) sum(length(which(is.na(y)))))

na_count_b
na_count_6m
na_count_1y
na_count_18m
```


```{r}
library(brms)

broom_data <- data %>%
  # select(QSEQQALY, time, AGE, SEX, VSBMI, LBESR, DMSMK, RFCTCAVI, MBSPECIES, CMQUIT_GROUP_2, SUBJNO) %>%
  select(QSEQQALY, time, AGE, SEX, VSBMI, DMSMK, MBSPECIES, CMQUIT_GROUP_2, SUBJNO) %>%
  drop_na() %>%
  mutate(censored = ifelse(QSEQQALY == 1, 1, 0))


# Formula for your multilevel, longitudinal Tobit model with control variables
formula <- bf(
  QSEQQALY | cens(censored) ~ 
    time * CMQUIT_GROUP_2 +  # Time by group interaction
    AGE + SEX + VSBMI + DMSMK +  # Control variables
    (1 + time | SUBJNO)  # Random intercept & slope for each patient
)

# Fit the model
tobit_model <- brm(
  formula = formula,
  data = broom_data,
  family = gaussian(),
  chains = 4,
  cores = 4,
  iter = 4000,
  warmup = 1000,
  control = list(adapt_delta = 0.95)  # Increases adaptation to avoid divergent transitions
)
```

```{r}
# Check model convergence
summary(tobit_model)
plot(tobit_model)

# Check trace plots for MCMC chains
mcmc_plot(tobit_model, type = "trace")

# Posterior predictive checks
pp_check(tobit_model)
```


```{r}
# Formula for your multilevel, longitudinal Tobit model with control variables
formula_simple <- bf(
  QSEQQALY | cens(censored) ~ 
    time * CMQUIT_GROUP_2 +  # Time by group interaction
    AGE + SEX + VSBMI + DMSMK +  # Control variables
    (1 | SUBJNO)  # Random intercept for each patient
)

# Fit the model
tobit_model_simple <- brm(
  formula = formula_simple,
  data = broom_data,
  family = gaussian(),
  chains = 4,
  cores = 4,
  iter = 4000,
  warmup = 1000,
  control = list(adapt_delta = 0.95)  # Increases adaptation to avoid divergent transitions
)
```

```{r}
# Check model convergence
summary(tobit_model_simple)
plot(tobit_model_simple)

# Check trace plots for MCMC chains
mcmc_plot(tobit_model_simple, type = "trace")
```


```{r}
# Posterior predictive checks
pp_check(tobit_model_simple)

# First, extract posterior predictive samples
posterior_samples <- posterior_predict(tobit_model_simple)

# Extract observed values (including censored values)
y_obs <- tobit_model_simple$data$QSEQQALY

# Create a custom posterior predictive check plot
# This requires ggplot2
library(ggplot2)
library(tidyverse)

# Sample a subset of posterior draws (e.g., 10 draws)
set.seed(123) # For reproducibility
n_draws <- 10
draw_indices <- sample(1:nrow(posterior_samples), n_draws)

# Create a data frame for plotting
plot_data <- data.frame(
  value = c(y_obs, as.vector(posterior_samples[draw_indices,])),
  type = c(rep("Observed", length(y_obs)), 
           rep("Replicated", length(y_obs) * n_draws)),
  draw = c(rep(NA, length(y_obs)), 
           rep(1:n_draws, each = length(y_obs)))
)

# Create the density plot
ggplot(plot_data, aes(x = value, color = type, group = interaction(type, draw))) +
  geom_density(data = subset(plot_data, type == "Replicated"), 
               aes(group = draw), size = 0.5, alpha = 0.3) +
  geom_density(data = subset(plot_data, type == "Observed"), 
               size = 1, alpha = 1) +
  scale_color_manual(values = c("Observed" = "navy", "Replicated" = "lightblue")) +
  labs(title = "Posterior Predictive Check",
       x = "QALY Value", y = "Density") +
  theme_minimal() +
  coord_cartesian(xlim = c(0, 1.3)) # Adjust as needed
```


Looking at above, we can see a couple of things:

1. Model Summary (Image 1)
model shows excellent convergence with all Rhat values at 1.00, indicating the chains have mixed well. 

The model successfully captured:
Group differences:
Groups B and C show different baseline QALY values compared to Group A (reference). 
Group B has an average of -0.03 lower QALY (95% CI: -0.09, 0.04) and 
Group C has -0.09 lower QALY (95% CI: -0.16, -0.02).

Time effects: There's a decreasing trend in QALY over time for Group A (reference group):
1Y: -0.01 (95% CI: -0.05, 0.03)
6M: -0.02 (95% CI: -0.06, 0.02)
B: -0.03 (95% CI: -0.06, 0.01)

Group × Time interactions: 
Group B shows positive interactions with time (suggests better trajectories compared to Group A):
B×Group B: 0.02 (95% CI: -0.03, 0.08)
6M×Group B: 0.06 (95% CI: -0.00, 0.12)
1Y×Group B: 0.05 (95% CI: -0.01, 0.11)

Group C shows a less dramatic, but overall negative interactions with time:
B x C: 0.02 (-0.04, 0.08)
6M x C: 0.03 (-0.03, 0.09)
1Y x C: 0.00 (-0.06, 0.06)

Control variables:
SEX2 has a significant negative effect (-0.15, 95% CI: -0.22, -0.07)
DMSMK1 has a significant negative effect (-0.12, 95% CI: -0.20, -0.03)
VSBMI has a small positive effect (0.01, 95% CI: 0.00, 0.02)
AGE has a negligible effect

Random effects: The standard deviation for the random intercept (sd(Intercept)) is 0.14, indicating substantial between-patient variability.
Residual error: Sigma is 0.11, which represents the residual variation after accounting for the fixed and random effects.

Trace Plots (Images 2-6)
The trace plots show excellent mixing of the MCMC chains:

All chains appear to explore the same parameter space with good overlap
No evidence of "stickiness" or autocorrelation
No trends or patterns are visible that would suggest non-convergence
The density plots (left side) show smooth, approximately normal posterior distributions
No evidence of multimodality or other convergence issues

overall - model converged well, and the posterior estimates are reliable.

Posterior Predictive Check (Image 7)
checking the effect of the right censoring at 1:

The observed data (dark blue line, 'y') shows a sharp peak near 1, demonstrating the censoring effect
The replicated data (light blue lines, 'y_rep') from the model don't capture this peak perfectly

This discrepancy is expected and validates the choice of a Tobit model. 
** The replicated data from the model represent what the uncensored data would look like, while the actual observed data show the censoring effect. This is precisely what the Tobit model is designed to handle.


Clinical Interpretation
From a clinical perspective, your model reveals several important findings:

Group C performs worse than the reference group (A) in terms of QALY scores.
Group B may have better trajectories over time than Group A, as evidenced by the positive interaction terms, though these are borderline significant.
Sex and smoking status are important covariates affecting QALY scores, with both showing negative associations.
BMI has a small positive association with QALY scores in this population.
Individual variation is substantial, as shown by the random intercept variance, highlighting the importance of the multilevel approach.

Statistical Validity
The model appears statistically sound with:
Good convergence (Rhat = 1.00 for all parameters)
Adequate effective sample sizes (all Bulk_ESS > 2000, Tail_ESS > 3500)
No divergences in the sampling
Appropriate handling of the censoring mechanism

```{r}
# Get fitted values for each patient at each time point
# This includes both fixed and random effects
fitted_values <- fitted(tobit_model_simple, summary = TRUE)

your_data <- broom_data

# Add these fitted values to your original data
your_data$fitted_qaly <- fitted_values[, "Estimate"]
# your_data$fitted_qaly_lower <- fitted_values[, "Q2.5"]
# your_data$fitted_qaly_upper <- fitted_values[, "Q97.5"]

# Calculate group-level statistics based on individual fitted values
# Calculate group summary statistics at each time point
group_summary <- your_data %>%
  mutate(time_continuous = case_when(
    time == "B" ~ 0,
    time == "6M" ~ 0.5,
    time == "1Y" ~ 1,
    time == "18M" ~ 1.5,
    TRUE ~ NA_real_  # Handle any unexpected values
  )) %>% 
  group_by(CMQUIT_GROUP_2, time_continuous) %>%
  summarize(
    mean_qaly = mean(fitted_qaly),
    q25_qaly = quantile(fitted_qaly, 0.25),
    q75_qaly = quantile(fitted_qaly, 0.75),
    sd_qaly = sd(fitted_qaly),
    n = n(),
    .groups = "drop"
  )

# Create the plot with group trajectories and quantile-based "confidence intervals"
group_colors <- c("A" = "#EEAAAA", "B" = "#AAAAEE", "C" = "#AAEEAA")

ggplot(group_summary, aes(x = time_continuous, y = mean_qaly, color = CMQUIT_GROUP_2, fill = CMQUIT_GROUP_2)) +
  geom_line(size = 1.5) +
  # Add ribbons for the 25th to 75th percentile range
  geom_ribbon(aes(x= time_continuous, ymin = q25_qaly, ymax = q75_qaly, fill = CMQUIT_GROUP_2), alpha = 0.2, colour = NA) +
  # x-axis should match time scale (B=0, 6M=0.5, 1Y=1, 18M=1.5)
  scale_x_continuous(
    breaks = c(0, 0.5, 1, 1.5),
    labels = c("B", "6M", "1Y", "18M")
  ) +
  labs(
    title = "QALY Trajectories by Treatment Group",
    subtitle = "Group Means with 25th-75th Percentile Ranges",
    x = "Time",
    y = "Quality-Adjusted Life Years (QALY)",
    color = "Group",
    fill = "Group"
  ) +
  theme_minimal() +
  scale_color_manual(values = group_colors) +
  scale_fill_manual(values = group_colors) +
  coord_cartesian(ylim = c(0.7, 1.1))
```

instead of the 25, 75th percentile values which display range of values, i want to use the Standard error of the mean (SEM) to visualize the uncertainty in the group mean.
```{r}
group_summary <- group_summary %>%
  mutate(
    # Create ±1 SD bounds (captures ~68% of data)
    lower_bound = pmax(0, mean_qaly - sd_qaly),  # Ensure lower bound is not below 0
    upper_bound = pmin(1, mean_qaly + sd_qaly)   # Ensure upper bound is not above 1
  ) %>%
  mutate(
    sem = sd_qaly / sqrt(n),
    lower_sem = pmax(0, mean_qaly - sem),
    upper_sem = pmin(1, mean_qaly + sem)
  ) %>% 
  mutate(
    # 95% CI using t-distribution
    margin = qt(0.975, n-1) * (sd_qaly / sqrt(n)),
    lower_ci = pmax(0, mean_qaly - margin),
    upper_ci = pmin(1, mean_qaly + margin)
  )

ggplot(group_summary, aes(x = time_continuous, y = mean_qaly, color = CMQUIT_GROUP_2, fill = CMQUIT_GROUP_2)) +
  geom_line(size = 1.5) +
  # mean ± 1 SD
  # geom_ribbon(aes(ymin = lower_bound, ymax = upper_bound), alpha = 0.2, colour = NA) +
  # SEM
  # geom_ribbon(aes(ymin = lower_sem, ymax = upper_sem), alpha = 0.2, colour = NA) +
  # 95% CI
  geom_ribbon(aes(ymin = lower_ci, ymax = upper_ci), alpha = 0.2, colour = NA) +
  scale_x_continuous(
    breaks = c(0, 0.5, 1, 1.5),
    labels = c("B", "6M", "1Y", "18M")
  ) +
  labs(
    title = "QALY Trajectories by Treatment Group",
    subtitle = "Group Means with 95% CI using t-distribution",
    x = "Time",
    y = "Quality-Adjusted Life Years (QALY)",
    color = "Group",
    fill = "Group",
    caption = "Shaded areas represent 95% CI, meaning we are 95% confidence that the true group mean falls within this range"
  ) +
  theme_minimal() +
  scale_color_manual(values = group_colors) +
  scale_fill_manual(values = group_colors) +
  coord_cartesian(ylim = c(0.7, 1.1))
```





```{r}
library(tidyverse)
library(brms)

# Assuming your model is called 'tobit_model'
# First, create a prediction grid with your time values
# I'll use your time variable structure, adjusting the approach

# Create a sequence of time points for each group
pred_data <- expand_grid(
  CMQUIT_GROUP_2 = c("A", "B", "C"),
  time = sort(unique(broom_data$time)),
  # Add mean values for your control variables
  AGE = mean(broom_data$AGE, na.rm = TRUE),
  SEX = names(which.max(table(broom_data$SEX))), # Most common sex
  VSBMI = mean(broom_data$VSBMI, na.rm = TRUE),
  DMSMK = names(which.max(table(broom_data$DMSMK))) # Most common smoking status
)

# Generate predictions from the model
# We use re_formula = NA to get population-level predictions (no random effects)
predictions <- fitted(
  tobit_model_simple, 
  newdata = pred_data,
  re_formula = NA,
  summary = TRUE
) %>% 
  as_tibble() %>%
  bind_cols(pred_data)

# Define your color scheme
group_colors <- c("A" = "#EEAAAA", "B" = "#AAAAEE", "C" = "#AAEEAA")

# Create the plot
tobit_plot <- ggplot(predictions, aes(x = time, y = Estimate, color = CMQUIT_GROUP_2)) +
  # Main regression lines
  geom_line(linewidth = 1) +
  # Confidence interval ribbons
  geom_ribbon(
    aes(ymin = Q2.5, ymax = Q97.5, fill = CMQUIT_GROUP_2),
    alpha = 0.2,
    linetype = 0  # No border for the ribbon
  ) +
  # Customize appearance
  labs(
    title = "Health-Related Quality of Life Trends by Treatment Response Group",
    x = "Time", 
    y = "QALY weight",
    color = "Response Group",
    fill = "Response Group"
  ) +
  scale_y_continuous(limits = c(0, 1)) +
  scale_fill_manual(values = group_colors) +
  scale_color_manual(values = group_colors) +
  theme_classic() +
  theme(
    legend.position = "bottom",
    plot.title = element_text(face = "bold", size = 14),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(size = 10)
  )

# Print the plot
print(tobit_plot)
```


```{r}
library(tidyverse)
library(AER)  # For tobit models
library(ggplot2)

# Handle censoring at 1.0
data <- data %>%
  mutate(QSEQQALY_adj = ifelse(QSEQQALY >= 1, 0.999, QSEQQALY))

# 2. Split data into the 12 groups
groups <- list()
for (t in c("B", "6M", "1Y", "18M")) {
  for (g in c("A", "B", "C")) {
    group_name <- paste0(t, "_", g)
    groups[[group_name]] <- data %>% 
      filter(time == t, CMQUIT_GROUP_2 == g)
    
    # Print sample size for this group
    cat("Group", group_name, "has", nrow(groups[[group_name]]), "observations\n")
  }
}

# 3. Variable selection for each group (optional, for simplicity)
# For example, using AIC-based stepwise selection
selected_vars <- list()
tobit_models <- list()

for (group_name in names(groups)) {
  # Get data for this group
  group_data <- groups[[group_name]]
  
  # Skip if sample size is too small
  if (nrow(group_data) < 10) {
    cat("Skipping group", group_name, "due to small sample size\n")
    next
  }
  
  # Full model with all variables
  full_formula <- as.formula("QSEQQALY_adj ~ AGE + SEX + VSBMI + LBESR + DMSMK + RFCTCAVI")
  
  # Fit initial tobit model
  full_model <- AER::tobit(full_formula, left = 0, right = 1, data = group_data)
  
  # Run step-wise selection (if sample size permits)
  if (nrow(group_data) >= 20) {
    step_model <- step(full_model, direction = "both", trace = FALSE)
    final_model <- step_model
    # Extract selected variables
    selected_vars[[group_name]] <- attr(terms(step_model), "term.labels")
  } else {
    # For very small samples, use a simpler model
    final_model <- tobit(QSEQQALY_adj ~ AGE + SEX, left = 0, right = 1, data = group_data)
    selected_vars[[group_name]] <- c("AGE", "SEX")
  }
  
  # Store final model
  tobit_models[[group_name]] <- final_model
  
  # Print summary
  cat("\n=== Tobit model for", group_name, "===\n")
  print(summary(final_model))
}

# 4. Visualize predicted values for each group
# Create a function to get predictions for a given value range
get_predictions <- function(model, data, variable, range) {
  # Create prediction grid
  pred_data <- data %>% 
    summarise(across(where(is.numeric), mean)) %>%
    slice(rep(1, length(range)))
  
  # Set the variable of interest to the range values
  pred_data[[variable]] <- range
  
  # Predict
  preds <- predict(model, newdata = pred_data, type = "response")
  
  # Return data frame
  return(data.frame(
    x = range,
    y = preds
  ))
}

# Example: Plot effect of AGE on QSEQQALY for each group at baseline
baseline_groups <- c("B_A", "B_B", "B_C")
age_range <- seq(min(data$AGE), max(data$AGE), length.out = 100)

age_effects <- data.frame()
for (group_name in baseline_groups) {
  if (!is.null(tobit_models[[group_name]])) {
    preds <- get_predictions(tobit_models[[group_name]], 
                            groups[[group_name]], 
                            "AGE", 
                            age_range)
    preds$group <- group_name
    age_effects <- rbind(age_effects, preds)
  }
}

# Plot age effects
ggplot(age_effects, aes(x = x, y = y, color = group)) +
  geom_line() +
  labs(title = "Effect of Age on QSEQQALY at Baseline by Group",
       x = "Age", y = "Predicted QSEQQALY") +
  theme_minimal()

# 5. Create summary table of coefficients across all 12 models
coef_table <- data.frame(
  Group = character(),
  Variable = character(),
  Coefficient = numeric(),
  StdError = numeric(),
  PValue = numeric(),
  stringsAsFactors = FALSE
)

for (group_name in names(tobit_models)) {
  model <- tobit_models[[group_name]]
  if (!is.null(model)) {
    coefs <- summary(model)$coefficients
    for (i in 1:nrow(coefs)) {
      coef_table <- rbind(coef_table, data.frame(
        Group = group_name,
        Variable = rownames(coefs)[i],
        Coefficient = coefs[i, 1],
        StdError = coefs[i, 2],
        PValue = coefs[i, 4]
      ))
    }
  }
}

# Print and save coefficient summary
print(coef_table)
write.csv(coef_table, "tobit_coefficients_by_group.csv", row.names = FALSE)

# 6. Generate predictions at representative values
# For example, predict QSEQQALY for a typical patient in each group
typical_patient <- data %>%
  summarize(across(where(is.numeric), median))

for (group_name in names(tobit_models)) {
  if (!is.null(tobit_models[[group_name]])) {
    pred <- predict(tobit_models[[group_name]], 
                  newdata = typical_patient, 
                  type = "response")
    cat("Predicted QSEQQALY for typical patient in group", 
        group_name, ":", pred, "\n")
  }
}
```

```{r}


# Assuming your data is loaded as 'data' with columns for:
# - patient_id: unique identifier for each patient
# - time: B, 6M, 12M, 18M
# - group: A, B, C
# - QSEQQALY: your outcome variable
# - Predictors: AGE, SEX, VSBMI, LBESR, DMSMK, RFCTCAVI

# 1. Create a list to store all models
all_models <- list()

# 2. Create a copy of the original data to add predictions
data_with_predictions <- data %>%
  select(QSEQQALY, time, AGE, SEX, VSBMI, LBESR, DMSMK, RFCTCAVI, MBSPECIES, CMQUIT_GROUP_2, SUBJNO)


na_count <- sapply(data_with_predictions %>% filter(time == "18M"), function(y) sum(length(which(is.na(y)))))
na_count

prediction_list <- list() 
# 3. Split data and apply tobit models
for (t in c("B", "6M", "1Y", "18M")) {
  for (g in c("A", "B", "C")) {
    # Create a subset for this time and group
    subset_name <- paste0(t, "_", g)
    
    # Extract data for this subset
    subset_data <- data %>% 
      filter(time == t, CMQUIT_GROUP_2 == g)
    
    # Skip if there's not enough data
    if (nrow(subset_data) < 10) {
      warning(paste("Skipping", subset_name, "due to insufficient data"))
      next
    }
    
    # Fit tobit model
    # Note: QSEQQALY is censored between 0 and 1
    model <- AER::tobit(
      QSEQQALY ~ AGE + SEX + VSBMI + LBESR + DMSMK + RFCTCAVI,
      left = 0, 
      right = 1, 
      data = subset_data
    )
    
    # Store the model
    all_models[[subset_name]] <- model
    
    # Generate predictions
    predictions <- predict(model, type = "response")
    
    # Store predictions in the list with identifiers
    prediction_df <- subset_data %>%
      select(SUBJNO, time, CMQUIT_GROUP_2) %>%
      mutate(predicted_QSEQQALY = predictions)
    
    prediction_list[[subset_name]] <- prediction_df
    
    # Print summary for this model
    cat("\n=== Model for", subset_name, "===\n")
    print(summary(model))
  }
}

# 4. Calculate model diagnostics for each subset
model_diagnostics <- data.frame(
  Model = character(),
  Sample_Size = numeric(),
  Sigma = numeric(),
  LogLik = numeric(),
  AIC = numeric(),
  BIC = numeric(),
  stringsAsFactors = FALSE
)

for (model_name in names(all_models)) {
  model <- all_models[[model_name]]
  
  model_diagnostics <- rbind(model_diagnostics, 
    data.frame(
      Model = model_name,
      Sample_Size = model$nobs,
      Sigma = model$sigma,
      LogLik = logLik(model),
      AIC = AIC(model),
      BIC = BIC(model),
      stringsAsFactors = FALSE
    )
  )
}

# Print diagnostics table
print(model_diagnostics)

# 5. Compare predictions to original values
data_with_predictions <- data_with_predictions %>%
  mutate(prediction_error = QSEQQALY - predicted_QSEQQALY)

# Calculate RMSE for each model
error_summary <- data_with_predictions %>%
  group_by(time, group) %>%
  summarize(
    RMSE = sqrt(mean(prediction_error^2, na.rm = TRUE)),
    MAE = mean(abs(prediction_error), na.rm = TRUE),
    Mean_Error = mean(prediction_error, na.rm = TRUE),
    .groups = 'drop'
  )

print(error_summary)

# 6. Visualize original vs predicted values
ggplot(data_with_predictions, aes(x = QSEQQALY, y = predicted_QSEQQALY, color = group)) +
  geom_point(alpha = 0.7) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed") +
  facet_grid(group ~ time, labeller = label_both) +
  labs(
    title = "Original vs Predicted QSEQQALY Values",
    x = "Original QSEQQALY",
    y = "Predicted QSEQQALY"
  ) +
  theme_minimal()

# 7. Visualize coefficient estimates across all models
# Extract coefficients from all models
coef_data <- data.frame()

for (model_name in names(all_models)) {
  model <- all_models[[model_name]]
  
  # Split the model name into time and group
  parts <- strsplit(model_name, "_")[[1]]
  time <- parts[1]
  group <- parts[2]
  
  # Extract coefficients
  coefs <- coef(model)
  
  # Add to dataframe
  for (var_name in names(coefs)) {
    coef_data <- rbind(coef_data, 
      data.frame(
        Time = time,
        Group = group,
        Variable = var_name,
        Coefficient = coefs[var_name],
        stringsAsFactors = FALSE
      )
    )
  }
}

# Plot coefficients
ggplot(coef_data %>% filter(Variable != "(Intercept)"), 
       aes(x = Variable, y = Coefficient, fill = Group)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_wrap(~ Time, ncol = 2) +
  labs(
    title = "Coefficient Estimates Across Time Points and Groups",
    x = "Variable",
    y = "Coefficient Estimate"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# 8. Export results
# Save the full dataset with predictions
write.csv(data_with_predictions, "tobit_predictions.csv", row.names = FALSE)

# Save model diagnostics
write.csv(model_diagnostics, "tobit_model_diagnostics.csv", row.names = FALSE)

# Save coefficient data
write.csv(coef_data, "tobit_coefficient_estimates.csv", row.names = FALSE)
```





model building
```{r}
# Start with all variables
full_model <- AER::tobit(QSEQQALY ~ AGE + SEX + DMSMK + RFCTCAVI + VSBMI + LBESR + CMQUIT_GROUP_2, 
                    left = 0, right = 1, 
                    data = subset(data, time == "B"))
summary(full_model)

# Check for multicollinearity
vif_values <- vif(lm(QSEQQALY ~ AGE + SEX + DMSMK + RFCTCAVI + VSBMI + LBESR + CMQUIT_GROUP_2, data = data))
print(vif_values)

# Backward selection based on AIC
step_model <- step(full_model, direction = "backward")
summary(step_model)

# Alternative approach: Create models with different subsets
model1 <- tobit(QSEQQALY ~ GROUP + AGE + SEX, left = 0, right = 1, data = data)
model2 <- tobit(QSEQQALY ~ GROUP + AGE + SEX + VSBMI, left = 0, right = 1, data = data)
model3 <- tobit(QSEQQALY ~ GROUP + AGE + SEX + VSBMI + LBESR, left = 0, right = 1, data = data)
model4 <- tobit(QSEQQALY ~ GROUP + AGE + SEX + DMSMK + RFCTCAVI + VSBMI + LBESR, left = 0, right = 1, data = data)

# Compare AIC
aic_values <- c(AIC(model1), AIC(model2), AIC(model3), AIC(model4))
models <- c("Group+Age+Sex", "Group+Age+Sex+BMI", "Group+Age+Sex+BMI+ESR", "All variables")
aic_comparison <- data.frame(Model = models, AIC = aic_values)
aic_comparison %>% arrange(AIC)

```

model transformation and alternatives
```{r}
# For heavily censored data, try a logit transformation
data$QSEQQALY_adj <- ifelse(data$QSEQQALY >= 1, 0.999, data$QSEQQALY)
data$QSEQQALY_logit <- log(data$QSEQQALY_adj/(1-data$QSEQQALY_adj))

# Fit tobit on transformed data
logit_model <- tobit(QSEQQALY_logit ~ GROUP + AGE + SEX + VSBMI + LBESR,
                     data = data)
summary(logit_model)

# Alternative: Beta regression for values between 0 and 1
# Requires the betareg package
library(betareg)
data$QSEQQALY_adj <- ifelse(data$QSEQQALY >= 1, 0.999, 
                           ifelse(data$QSEQQALY <= 0, 0.001, data$QSEQQALY))
beta_model <- betareg(QSEQQALY_adj ~ GROUP + AGE + SEX + VSBMI + LBESR, data = data)
summary(beta_model)

# Alternative: Fractional logit (using GLM with logit link)
frac_logit <- glm(QSEQQALY ~ GROUP + AGE + SEX + VSBMI + LBESR, 
                  family = quasibinomial(link = "logit"), 
                  data = data)
summary(frac_logit)
```

validation
```{r}
# Create training and test sets
set.seed(123)
train_indices <- sample(1:nrow(data), 0.7 * nrow(data))
train_data <- data[train_indices, ]
test_data <- data[-train_indices, ]

# Fit model on training data
train_model <- tobit(QSEQQALY ~ GROUP + AGE + SEX + VSBMI + LBESR, 
                     left = 0, right = 1, 
                     data = train_data)

# Predict on test data
# Need to handle the censoring in predictions
predictions <- predict(train_model, newdata = test_data, type = "response")

# Calculate performance metrics
rmse <- sqrt(mean((predictions - test_data$QSEQQALY)^2, na.rm = TRUE))
mae <- mean(abs(predictions - test_data$QSEQQALY), na.rm = TRUE)
print(paste("RMSE:", rmse))
print(paste("MAE:", mae))

# Cross-validation 
library(caret)
train_control <- trainControl(method = "cv", number = 5)
cv_model <- train(QSEQQALY ~ GROUP + AGE + SEX + VSBMI + LBESR, 
                 data = data,
                 method = "lm",  # Approximation since tobit isn't directly supported
                 trControl = train_control)
print(cv_model)
```

interpretation
```{r}
# Extract coefficients and confidence intervals
coef_table <- summary(step_model)$coefficients
conf_int <- confint(step_model)
coef_data <- data.frame(
  Variable = rownames(coef_table),
  Estimate = coef_table[, 1],
  StdError = coef_table[, 2],
  LowerCI = conf_int[, 1],
  UpperCI = conf_int[, 2],
  PValue = coef_table[, 4]
)

# Calculate marginal effects (average partial effects)
library(margins)
margins_model <- margins(step_model)
summary(margins_model)

# Visualize coefficient effects
ggplot(coef_data[-1,], aes(x = reorder(Variable, Estimate), y = Estimate)) +
  geom_point() +
  geom_errorbar(aes(ymin = LowerCI, ymax = UpperCI), width = 0.2) +
  coord_flip() +
  labs(title = "Coefficient Estimates with 95% Confidence Intervals",
       x = "Variable", y = "Estimate")

# Visualize predicted QSEQQALY by GROUP, holding other variables at their means
new_data <- expand.grid(
  GROUP = unique(data$GROUP),
  AGE = mean(data$AGE),
  SEX = factor(levels(data$SEX)[1], levels = levels(data$SEX)),
  VSBMI = mean(data$VSBMI),
  LBESR = mean(data$LBESR)
)
new_data$predicted <- predict(step_model, newdata = new_data, type = "response")

ggplot(new_data, aes(x = GROUP, y = predicted, fill = GROUP)) +
  geom_col() +
  labs(title = "Predicted QSEQQALY by Group")
```

stratify
```{r}
# For improved insights, consider stratifying by GROUP
group_levels <- unique(data$GROUP)
for (g in group_levels) {
  group_data <- data %>% filter(GROUP == g)
  group_model <- tobit(QSEQQALY ~ AGE + SEX + VSBMI + LBESR, 
                      left = 0, right = 1, 
                      data = group_data)
  cat("\nResults for Group", g, ":\n")
  print(summary(group_model))
}
```

longitinal
```{r}
# If you have longitudinal data
library(nlme)
library(lme4)

# Approximate with linear mixed effects model
lmm_model <- lmer(QSEQQALY ~ GROUP + AGE + SEX + VSBMI + LBESR + (1|SUBJECT_ID), 
                 data = data)
summary(lmm_model)

# Or using a specific longitudinal tobit package if available
# This is just a placeholder; implementation depends on specific packages
# long_tobit <- long_tobit_function(QSEQQALY ~ GROUP + AGE + SEX + VSBMI + LBESR, 
#                                  subject = "SUBJECT_ID", 
#                                  data = data)
```

final
```{r}
# Select the best model based on AIC, BIC, and theoretical considerations
final_model <- step_model  # Or whichever model you decide is best

# Create a summary table for reporting
final_summary <- data.frame(
  Variable = rownames(summary(final_model)$coefficients),
  Coefficient = summary(final_model)$coefficients[, 1],
  StdError = summary(final_model)$coefficients[, 2],
  PValue = summary(final_model)$coefficients[, 4],
  Significance = ifelse(summary(final_model)$coefficients[, 4] < 0.05, "*", "")
)

# Export results
write.csv(final_summary, "tobit_model_results.csv", row.names = FALSE)

# Final visualizations for reporting
# Observed vs Predicted
data$predicted <- predict(final_model, type = "response")
ggplot(data, aes(x = predicted, y = QSEQQALY)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1, color = "red") +
  labs(title = "Observed vs Predicted QSEQQALY Values")
```

